---
title: "Precipitation Spatial Analysis"
author: "Tarun Reddy Aleti"
date: "April 18, 2016"
output: html_document
---

### Executive Summary
- Using NOAA free open datasets trying to implement preliminary spatial data analysis using R
- We will be using Precipitaion data [ftp://ftp.ncdc.noaa.gov/pub/data/normals/1981-2010/source-datasets/]
- The data set consists of monthly average precipitation at every station every month from 1981 to 2010.
- Whenever a station recording is missing then its replaced by -9999

### Required Packages
```{r, echo = TRUE, warning = FALSE, message = FALSE, results = 'hide'}
library(sp)
library(maptools)
library(maps)
library(rgdal)
library(lattice)
library(grid)
library(RColorBrewer)
library(googleVis)
library(RgoogleMaps)
library(sqldf)
library(rworldmap)
library(dismo)
```

### Data
```{r, echo = TRUE, warning = FALSE, message = FALSE, results = 'hide'}
stations <- read.fwf("ghcnd-stations.txt", widths = c(11, 9, 10, 7, 3, 31, 4, 4, 6, 14))
colnames(stations) <- c("ID", "LATITUDE", "LONGITUDE", "ELEVATION", "STATE", "NAME",
                        "GSNFLAG", "HCNFLAG", "WMOID", "METHOD")

prcp <- read.fwf("mly-prcp-filled.txt", widths = c(11, 5,  6, 2, rep(c(5, 2), 11)))
colnames(prcp) <- c("STNID", "YEAR", paste(c("VALUE", "FLAG"), rep(c(1:12), each = 2), sep = ""))

data <- sqldf("SELECT s.*, p.*
               FROM stations as s
               INNER JOIN prcp as p on s.ID = p.STNID")
```

### Spatial Points and Graphics

- For our analysis we shall be using data from Jan 2000 in Texas state

```{r, echo = TRUE, warning = FALSE, message = FALSE}
prcp_texas <- subset(data, data$STATE == " TX" & data$YEAR == "2000")
prcp_texas <- prcp_texas[ ,c("LATITUDE", "LONGITUDE", "VALUE1")]
prcp_texas$LATITUDE <- as.numeric(prcp_texas$LATITUDE)
prcp_texas$LONGITUDE <- as.numeric(prcp_texas$LONGITUDE)
coordinates(prcp_texas) <- c("LONGITUDE", "LATITUDE")
crs.geo <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84")
proj4string(prcp_texas) <- crs.geo
summary(prcp_texas)
```

- Quickly plotting point data on a map

```{r, echo = TRUE, warning = FALSE, message = FALSE}
plot(prcp_texas, pch = 20, col = "steelblue")
plot(coastsCoarse, add = T)
```

- Mapping vectorial data using gmap from dismo

```{r, echo = TRUE, warning = FALSE, message = FALSE}
prcp_texas_gbmap <- gmap(prcp_texas, type = "satellite")
prcp_texas_merc <- Mercator(prcp_texas)
plot(prcp_texas_gbmap)
points(prcp_texas_merc, pch = 20, col = "red")

prcp_texas_coords <- as.data.frame(coordinates(prcp_texas))  # retrieves coordinates 

PlotOnStaticMap(lat = prcp_texas_coords$LATITUDE, lon = prcp_texas_coords$LONGITUDE, zoom = 5, cex = 1.4, pch = 19, col = "red", FUN = points,  size = c(640, 640), add = F)

map.lim <- qbbox(prcp_texas_coords$LATITUDE, prcp_texas_coords$LONGITUDE, TYPE = "all") 

mymap <- GetMap.bbox(map.lim$lonR, map.lim$latR, destfile = "gmap.png", maptype = "satellite")

PlotOnStaticMap(mymap, lat = prcp_texas_coords$LATITUDE, lon = prcp_texas_coords$LONGITUDE, zoom = NULL, cex = 1.3, pch = 19, col = "red", FUN = points, add = F)
```

- Drawing Polygons and Polylines

```{r, echo = TRUE, warning = FALSE, message = FALSE}
#plot(prcp_texas_gbmap)
#mypolygon <- drawPoly()
#summary(mypolygon)
```

### Converting between formats, reading in, and saving spatial vector data

#### Exporting KML (Google Earth)
```{r, echo = TRUE, warning = FALSE, message = FALSE}
writeOGR(prcp_texas, dsn = "locsgb.kml", layer = "locs2.gb", driver = "KML")
```

#### Reading KML
- Save as Shapefile
```{r, echo = TRUE, warning = FALSE, message = FALSE}
newmap <- readOGR("locsgb.kml", layer = "locs2.gb")
writePointsShape(prcp_texas, "locsgb")
```

- Read as Shapefile
```{r, echo = TRUE, warning = FALSE, message = FALSE}
gb.shape <- readShapePoints("locsgb.shp")
plot(gb.shape)
```


### Changing projection of spatial vector data

- Projecting point dataset: To illustrate, let's project the dataframe with coordinates that we obtained above:
```{r, echo = TRUE, warning = FALSE, message = FALSE}
summary(prcp_texas)
```

- The original coordinates are in lat lon format. Let's define the new desired projection: Lambert Azimuthal Equal Area in this case (look up parameters at http://spatialreference.org)

```{r, echo = TRUE, warning = FALSE, message = FALSE}
crs.laea <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")
prcp_texas_laea <- spTransform(prcp_texas, crs.laea)  # spTransform makes the projection
```

#### Projecting shapefile of countries

```{r, echo = TRUE, warning = FALSE, message = FALSE}
plot(countriesLow)  # countries map in geographical projection
country.laea <- spTransform(countriesLow, crs.laea)
plot(prcp_texas_laea, pch = 20, col = "steelblue")
plot(country.laea, add = T)
```

- Define spatial limits for plotting
```{r, echo = TRUE, warning = FALSE, message = FALSE}
plot(prcp_texas_laea, pch = 20, col = "steelblue", xlim = c(1800000, 3900000), ylim = c(1e+06, 
    3e+06))
plot(country.laea, add = T)
```
